---
title       : "Week 3: Introduction to the Tidyverse"
subtitle    : '`r format(Sys.Date(), "%m/%d/%Y")`'
author      : "Jake Campbell"
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

```{r, results='hide', echo=F, message=F}
library(tidyverse)

initial.tb<- 
  table4a %>%
  rename(Year1=`1999`, Year2=`2000`)

initial.starwars<- starwars[, 1:10]
```


## What is the Tidyverse?

- The tidyverse is a set of packages developed for data science and manipulation
    + They share a common philosophy and grammar
    + Intended to make code more understandable
    
- R is a bit antiquated

- Several different people contributing code can lead to confusion and overlapping ideas

- Tidyverse packages have different purposes, but read easily into each other

--- 

## Download the Packages

- The tidyverse consists of several different packages

- We can download them all at once, though

 + Just call `install.packages("tidyverse")`

---

## Reshaping Data

- Data generally comes in wide and long format

- Wide format makes every value a column

- Long format puts multiple values in a single column, creating another column that provides a key to the value column 

- Consider the goal of your analysis when choosing which format

---

## Reshaping with `tidyr`

- `tidyr` allows us to easily go from long to wide or vice-versa

- The main functions we'll use are `gather()` and `spread()`
  + These change datasets into long and wide formats, respectively

---

## Going from Wide to Long

- We'll use `gather()` to go from wide to long
- We need to first specify what columns we are gathering
  + The names of the columns will become our key, their values become our value
  + We need to also specify the new `key` and `value` column names
  
```{r}
gather(Year1, Year2, data = initial.tb, key = Year, value = Cases)
```


---

## Data Manipulation with `dplyr`

- `dplyr` has several functions for manipulating data frames

- We can chain functions together using `%>%`
  + Also known as the pipe operator
  + Note that if for some reason we needed to refer back to the tibble we are manipulating, we would specify it with a `.`
  
```{r, eval=F}
# Format would look like this

data_frame_x %>%
  function_x()
```


---

## Some Common `dplyr` Functions

- `select()`: pick certain variables
- `filter()`: filter your data on a given condition
- `arrange()`: order your data
- `mutate()`: create a new column
- `rename()`: rename your columns

---

## `select`

- With `select`, we just specify the variables we want to choose
  + We can instead specify what columns we don't want by putting a `-` before the column
  
```{r}
initial.starwars %>%
  # Select height and birth year of all characters
  select(height, birth_year) %>%
  head()
```

---

## `filter`

- With `filter`, we subset the tibble on some condition
  + Remember that `==` means equal to and `!=` means not equal to
  
```{r}
initial.starwars %>%
  # Filter on only males
  filter(gender == "male") %>%
  head()
```

---

## `arrange`

- With `arrange`, we order the tibble by some column
  + By default, order is ascending; to get descending order, we must use `desc()`
  
```{r}
initial.starwars %>%
  # Arrange by descending height
  arrange(desc(height)) %>%
  head()
```

---

## `mutate`

- `mutate` allows us to create new columns
  + Remember to save your output if you want to keep it!
  
```{r}
new.starwars<-
  initial.starwars %>%
  # Creating a new column that is the square root of height
  mutate(sqrt_height = sqrt(height))

new.starwars %>%
  select(height, sqrt_height)
  head(3)
```

---

## `rename`

- `rename` allows us to change the name of a column
  + First specify the new name, than specify what column is being renamed

```{r}
initial.starwars %>%
  # Rename name to full_name
  rename(full_name = name) %>%
  head()
```

---

## `summarize`

- With `summarize`, we can perform certain summary statistics, like finding the average, or getting a count

```{r}
initial.starwars %>%
  # Here we are finding the average height of all characters
  summarize(avg.height = mean(height, na.rm=T))
```

---

## `group_by`

- `group_by` allows us to perform different operations by groups
  + If we want to end the grouping, use `ungroup()`
  
```{r}
initial.starwars %>%
  # Grouping by gender
  group_by(gender) %>%
  # Finding average height
  summarize(avg.height = mean(height, na.rm=T)) %>%
  # Remember to ungroup if you want to perform non-grouped functions 
  # after grouping
  ungroup()
```

---

## Chaining Functions

- With the `%>%` operator, we can easily chain several functions together

```{r}
initial.starwars %>%
  # Filter characters taller than 200 cm
  filter(height > 200) %>%
  # Select the mass column
  select(mass) %>%
  # Order mass in descending order
  arrange(desc(mass))
```
