---
title       : "Week 4: The Grammar of Graphics"
subtitle    : '`r format(Sys.Date(), "%m/%d/%Y")`'
author      : "Jake Campbell"
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

```{r, results='hide', echo=F, message=F, warning=F}
library(tidyverse)

data("diamonds")
set.seed(1234)
diamonds <- diamonds %>%
  sample_n(size = 5000, replace = F)
```

## Identifying Relationships Graphically

- The most straightforward method for identifying relationships among variables is graphically
- R has several plotting functions in its base code
  + Quick to access, but hard to customize and not very visually appealing
- `ggplot2` is an outside plotting package that is part of the tidyverse
  + Similar grammar and ideas to what we've seen with `dplyr`
+ Great for publication-ready graphs

---

## The Grammar of Graphics

- Think of a graph as a sentence
- A sentence has several different aspects (noun, verb, adjective, etc.)
- A graph can be thought of as doing the same thing (data, geom, scale, etc.)
  + `ggplot` allows us to plot in this grammar of graphics style
  + Each part is connected by a `+`
  + Similar to how we connect each line of code in `dplyr` with a `%>%`

---

## Data

- The first argument in our ggplot code signifies where the data is coming from
- Specific aesthetics about the data (i.e. how data is mapped)
  + x and y values
  + Colors, fills, shapes

---

## Aesthetics

- If we want to map a variable to an aspect of our plot, we can include it in `aes()`
```{r, eval=F}
# Map gender to color
ggplot(aex(x = height, y = weight, col = gender))
```

- If we just want to map a single characteristic, we should do it outside of the `aes()` call
```{r, eval=F}
# Map the color blue to color
ggplot(aes(x = height, y = weight), col = "blue")
```

---

## Geom Object

- How we choose to visually represent our data
- Anything from a scatter plot to a text label
- Geoms can have their own attributes or aesthetics
  + i.e. color, transparency, size, etc.
- Geoms can easily be layered on top of each other

---

## Scatter Plots

- We use `geom_point()` to specify a scatterplot

```{r, eval=F}
ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point()
```

---

## Scatter Plots

```{r, echo=F, fig.align='center'}
ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point()
```

---

## Histograms

- We use `geom_histogram()` to create a histogram

```{r, eval=F, message=F}
ggplot(data = diamonds, aes(x = price)) +
  geom_histogram()
```

---

## Histograms

```{r, echo=F, fig.align='center', message=F}
ggplot(data = diamonds, aes(x = price)) +
  geom_histogram()
```

---

## Density Plots

- We use `geom_density()` to create a density plot

```{r, eval=F}
ggplot(data = diamonds, aes(x = price)) +
  geom_density()
```

---

## Density Plots

```{r, echo=F, fig.align='center'}
ggplot(data = diamonds, aes(x = price)) +
  geom_density()
```

---

## Barplots

- We use `geom_bar()` to create a barplot
- Could use the `position` argument to change bar placements
  + `"dodge"` creates side-by-side bars for groups
  + `"stack"` creates a stacked bar chart for groups
  + `"fill"` stacks the groups using ratios instead of counts

```{r, eval=F}
ggplot(data = diamonds, aes(x = cut)) +
  geom_bar()
```

---

## Barplots

```{r, echo=F, fig.align='center'}
ggplot(data = diamonds, aes(x = cut)) +
  geom_bar()
```

---

## Boxplots

- We can plot boxplots with `geom_boxplot()`

```{r, eval=F}
ggplot(data = diamonds, aes(x = cut, y = price)) +
  geom_boxplot()
```

---

## Boxplots

```{r, echo=F, fig.align='center'}
ggplot(data = diamonds, aes(x = cut, y = price)) +
  geom_boxplot()
```

---

## Faceting

- What if we want to look at the same x-y relationship for different groups or factor levels?
  + ex: the relationship between pain measure and time after surgery by gender
- Faceting allows us to do this
- `facet_wrap()` takes a look at our data by group
  + `facet_grid()` allows us to do this with multiple dimensions

---

## Scale Functions

- There are several scale functions that allow us to edit scales to either better represent our data, or customize our graph (notice functions I write here can be applied to the y axis by substituting x for y)
  + `scale_x_discrete()`: allows us to edit scale info on categorical data
  + `scale_x_continuous()`: allows us to edit scale info on numeric data
  + `scale_color_.`: can be used to edit scale info for data that has been mapped to color
    + Note, you can specify discrete, continuous, or manual
    + Also commonly applied to fill with `scale_fill_continuous` or `scale_fill_manual`

---

## Additional Components

- Graph labels can be edited using the `labs()` argument
  + x, y, title, fill, color, etc.
- `theme()` can be used to make changes to non-data aspects of the graph
  + Font type, color, legend position, etc.

