---
title       : "Week 2: Data Objects"
subtitle    : '`r format(Sys.Date(), "%m/%d/%Y")`'
author      : "Jake Campbell"
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Data Types

- Data can take on many different types of values
  + Data is not just numbers!
    + What about text, images, etc.?

--- 

## Major Data Types

- Numeric: number values
```{r}
class(100)
```
- Character: text strings specified by quotation marks
```{r}
class("Jake")
```

---

## Major Data Types

- Factor: categorical data
```{r}
class(factor(c("second", "first"), levels = c("first", "second")))
```
- Logical: boolean true or false
  + Specified either with all capital letters (`TRUE`) or by just the first letter (`T`)
```{r}
class(FALSE)
```

---

## Factors

- Factors are for categorical data rather than just any string data
- Categorical data should have a limited amount of possible options
  + There are only so many months in a year, but an infinite number of names
```{r}
test_factor <- factor(c("second", "first"), levels = c("first", "second"))
test_factor

as.numeric(test_factor)
```

---

## Factors

- Remember that factors are DIFFERENT than characters
- If we want to perform text manipulation, we need our data to be in character format
- A lot of models in R require data to be in factor form rather than character

---

## Changing Between Data Types

- To change between data types, we specify the data type we want to change to, prefaced by `as.`
```{r}
# Changing from numeric to character
test_character <- as.character(100)
class(test_character)

# Changing from character to numeric
test_number <- as.numeric(test_character)
class(test_number)
```

---

## Changing Between Data Types

- The only hiccup you might experience is going from factor to numeric
- Calling `as.numeric` on a factor will give you the level order rather than the expected labels
```{r}
test_factor <- factor(1:3, levels = c(2, 3, 1))
test_factor
as.numeric(test_factor)
```

---

## Changing Between Data Types

- We would instead need to wrap the factor in `as.character` first before `as.numeric`
```{r}
as.numeric(as.character(test_factor))
```

---

## Data Objects

- So we know data types, but what do we store the data in???
- R has different types of data objects
  + Each have their own requirements and presentations of data

---

## Vectors

- Vectors are just a simple combination of data of the same type
  + `c` is the combine function
```{r}
test_vector <- c(1:10)
test_vector
```

---

## Matrices

- Matrices can hold two dimensions of data of the same type
  + Rows and columns
- Computationally fast and usable for matrix applications
```{r}
# We have to specify the number of rows and columns in a matrix
test_matrix <- matrix(data = 1:9, nrow = 3, ncol = 3)
test_matrix
```

---

## Lists

- Lists can contain data of different types
- They can be a bit confusing to work with
```{r}
test_list <- list(name = c("Jake", "Hannah", "Jim"), 
                  age = c(25, 26, 33))
test_list
```

---

## Data Frames 

- Data frames are going to be the type of object we work with most
- Multi-dimensional, like a matrix, but can hold multiple data types
- Resembles a spreadsheet in appearance
```{r}
test_df <- data.frame(name = c("Jake", "Hannah", "Jim"), 
                      age = c(25, 26, 33))
test_df
```

---

## Indexing Objects

- What if we want the 3rd observation from a vector, or the 5th column of a dataframe?
  + We can index these in multiple ways
- Brackets allow us to specify element, row, or column of a data object
```{r}
test_vector[3]
```

---

## Indexing Objects

- Indexing the data frame requires a comma because it has two dimensions 
  + Put a 1 in front of the comma and it returns the first row
  + Put a 1 after the comma and it returns the first column
```{r}
test_df[1, ]
test_df[, 1]
```

---

## Indexing Objects

- Certain objects can also be indexed by the $ followed by the column name we want to select
```{r}
test_df$age
```

---

## Function Syntax

- We can write our own functions to use later on in R
- We first specify the `function()` function
  + Its arguments are the names of the arguments we’ll include in our function
- Next, we create curly braces `{}`
  + Everything inside our curly braces is the body of the function
  + What is saved in the curly braces isn’t saved in the global environment
```{r}
test_function <- function(input) {
  input + 1
}

test_function(1)
```

---

## Packages

- R has a huge library of packages that expand on base R functions
- Some packages are installed with your installation of R
  + We can load these with the `library()` function
```{r}
library(MASS)
```
- Most packages we'll use need to be installed first using `install.packages()`
- We only need to install a package once, but we need to load it up every new R session