---
title       : "Week 5: The Importance of Exploratory Analysis"
subtitle    : '`r format(Sys.Date(), "%m/%d/%Y")`'
author      : "Jake Campbell"
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : mathjax            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

```{r, results='hide', echo=F, message=F, warning=F}
library(tidyverse)
library(GGally)

data(iris)
```

## What is Exploratory Analysis?

- Exploratory analysis is an initial investigative approach to our data prior to more complex analysis

- Helps us get a good understanding of any general patterns in our data

- Very visual-based
  + Lots of plotting!

---

## Why is Exploratory Analysis Important

- Machine learning models are great, but they do not have the eye that humans do
  + An important variable to the model may just be the result of an error in our data that we would've captured with exploratory analysis
  
- By doing some initial investigation with our data, we can build better models down the road

- Important to get a feel for our data
  + You may not have a ton of knowledge about the topic you're analyzing, but it's important to build some understanding!

---

## Summary Statistics

- Summary statistics give us a quick overview of different aspects of our data (how spread out it is, what's the typical value, etc.)

- Good for quick insight, but don't always tell the whole story
  + Ex: `mean` of a skewed variable

---

## Common Summary Statistics

- The mean is the average of a vector of values
  + `R` function is `mean()`

- The median is the middle value in a vector of values
  + `R` function is `median()`
  
```{r}
mean(iris$Sepal.Length)

median(iris$Sepal.Length)
```

---

## Common Summary Statistics

- Quantiles divide observations into bins
  + The median is a quantile (splits the data in two)
  + `R` function is `quantile()`

```{r}
quantile(iris$Sepal.Length)

# Specify where the splits are
quantile(iris$Sepal.Length, probs = c(0, .33, .66, 1))
```

---

## Common Summary Statistics

- Variance is the average squared difference of a set of observations from the mean
  + $\frac{\sum (x_{i} - mean(x))^2}{n-1}$
  + Commonly represented as $\sigma^2$
  + `R` function is `var()`
  
- Standard deviation is square root of the variance
  + Commonly represented as $\sigma$
  + `R` function is `sd()`
  + Puts variance measure on the level of the data

---

## Common Summary Statistics

```{r}
var(iris$Sepal.Length)

sd(iris$Sepal.Length)
```

---

## Common Summary Statistics

- Covariance combines the variance of two variables, `x` and `y` and says whether they have a positive or negative relationship
  + $\frac{\sum (x_{i} - mean(x))(y_{i} - mean(y))}{n-1}$
  + `R` function is `cov()`

- Correlation divides the covariance of `x` and `y` by the standard deviations of `x` and `y`
  + $\frac{cov(xy)}{sd(x)sd(y)}$
  + Puts the covariance on a scale from `-1` to `1`
  + `R` function is `cor()`

---

## Common Summary Statistics

```{r}
cov(iris$Sepal.Length, iris$Sepal.Width)

cor(iris$Sepal.Length, iris$Sepal.Width)
```

---

## Visualizing Our Data

- The best method for us to investigate our data is visually

- We can easily identify patterns or notice issues with our data that we would miss otherwise

- It's important that we use plots correctly!

---

## When to Use What Plot

- Scatterplots are great for showing the relationship between continuous variables

- Histograms and density plots help us visualize the distribution of a variable

- Boxplots can show the relationship between a continuous and categorical variable

- Barplots are good for giving count data of a categorical variable

---

## Common Plot No-no's

- Don't mess with the limits of your plot

- This can easily be manipulated to show relationships that aren't there

```{r, echo=F, fig.align='center', fig.height=5}
iris %>%
  .[1:145, ] %>% 
  ggplot(aes(x = Species)) +
  geom_bar() +
  coord_cartesian(ylim = c(40, 50))
```

---

## Common Plot No-no's

- Avoid pie charts

- Although very popular, hard to interpret exact differences between groups

- Much easier to visualize in a bar chart

---

## Creating Multiple Plots at Once

- We can use the `GGally` package to create several extensions off of `ggplot2`

- One of these extensions is a plot matrix
  + Creates a matrix of plots, each showing a relationship between different variables
  + Function to use is `ggpairs`

---

## `ggpairs`

```{r, message=F, fig.align='center'}
ggpairs(iris)
```